---
title: "Assignment 5: Data Visualization"
author: "Xueying Feng"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.
8. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Salk_A05_DataVisualization.Rmd") prior to submission.

The completed exercise is due on Tuesday, February 11 at 1:00 pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse and cowplot packages. Upload the NTL-LTER processed data files for nutrients and chemistry/physics for Peter and Paul Lakes (tidy and gathered) and the processed data file for the Niwot Ridge litter dataset. 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1
getwd()

library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggridges)

PeterPaul.chem <- 
  read.csv("./Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv")
PeterPaul.gathered <-
  read.csv("./Data/Processed/NTL-LTER_Lake_Nutrients_PeterPaulGathered_Processed.csv")
Litter <- read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv")

#2
PeterPaul.chem$sampledate <- as.Date(PeterPaul.chem$sampledate, format = "%Y-%m-%d")
PeterPaul.gathered$sampledate <- as.Date(PeterPaul.gathered$sampledate, format = "%Y-%m-%d")
Litter$collectDate <- as.Date(Litter$collectDate, format = "%Y-%m-%d")

```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
mytheme <- theme_minimal(base_size = 12, base_family = "Times") +
  theme(axis.text.x = element_text(color = "DarkGrey"), 
        legend.position = "top")

theme_set(mytheme)

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best practices for data visualization. Ensure your theme, color palettes, axes, and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus by phosphate, with separate aesthetics for Peter and Paul lakes. Add a line of best fit and color it black. Adjust your axes to hide extreme values.

```{r}
install.packages("viridis")
library(viridis)

NvsP <-
  ggplot(PeterPaul.chem, aes(x = po4, y = tp_ug, color = depth, shape = lakename)) +
  geom_point(alpha = 0.95, size = 2) +  
  labs(x=expression(paste("Phosphate (",mu,"g/L)"))) + 
  labs(y=expression(paste("Total Phosphorus (",mu,"g/L)")))+
  labs(color="Depth(m)")+ 
  scale_shape_manual(values = c(16,17)) +  
  scale_color_viridis(option = "A", direction = -1) + 
  theme(legend.position = "right", 
        legend.text = element_text(size = 10), legend.title = element_text(size = 13))+
  geom_smooth(method = "lm", colour="black", size=0.5, se = FALSE)+ # default is two curve lines;
                                                                     #linear model also have two lines
  xlim(0, 50) +
  ylim(0, 150)

print(NvsP)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and (c) TN, with month as the x axis and lake as a color aesthetic. Then, create a cowplot that combines the three graphs. Make sure that only one legend is present and that graph axes are aligned.

```{r}
Nutrientplot <-
  ggplot(PeterPaul.chem, aes(x = lakename, y = concentration, color = nutrient)) +
  geom_boxplot() +
  labs(x=expression(paste("concentration (mg/L"))) + # change your legend labels here
  labs(y=expression(paste(".")))+
  scale_color_manual(labels = c("NH34", "NO23","PO4","TN","YP"),
                     values = c("#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84"))+
# place your additional edits here
  scale_y_continuous(expand = c(0, 0)) +
  #scale_color_brewer(palette = "YlGnBu") +
  #scale_color_manual(values = c("#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")) +
  #scale_color_viridis(discrete = TRUE, end = 0.8) +
  theme(legend.position = "right")
print(Nutrientplot)

```

Question: What do you observe about the variables of interest over seasons and between lakes? 

> Answer: 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the "Needles" functional group. Plot the dry mass of needle litter by date and separate by NLCD class with a color aesthetic. (no need to adjust the name of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into three facets rather than separated by color. 

```{r}



```

Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: 