---
title: "Assignment 2: Coding Basics"
author: "Xueying Feng"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.
8. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Salk_A02_CodingBasics.Rmd") prior to submission.

The completed exercise is due on Tuesday, January 21 at 1:00 pm.


## Basics Day 1

1. Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name. 

2. Compute the mean and median of this sequence. 

3. Ask R to determine whether the mean is greater than the median. 

4. Insert comments in your code to describe what you are doing.

```{r}
#1. 
Fours_1_100<-seq(1, 100, 4) # from 1 to 100, by 4
Fours_1_100 # renamed the sequence


#2. 
mean(Fours_1_100) 
median(Fours_1_100)

#3. 
mean(Fours_1_100) > median(Fours_1_100)
mean(Fours_1_100) == median(Fours_1_100)
# According to the results, mean is not greater than the median. They are equal.

```

## Basics Day 2

5. Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6. Label each vector with a comment on what type of vector it is.

7. Combine each of the vectors into a data frame. Assign the data frame an informative name.

8. Label the columns of your data frame with informative titles.

```{r}
#a
Names<-c("Lily","Bill","Kiki","Elsa") # character vector
Names
Scores<-c(80,70,95,45) # numeric vector
Scores
Name_Score <- data.frame(Names,Scores)
Name_Score

Pass <- function(Scores){
  ifelse(Scores>50, print("true"),print("false"))
}

Score1 <- Pass(80); 
Score2 <- Pass(70); 
Score3 <- Pass(95); 
Score4 <- Pass(45); 

```

9. QUESTION: How is this data frame different from a matrix?

> Answer: In a data frame, differnet columns can contain different types of data, but in a matrix, they are the same type of data

10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.

11. Apply your function to the vector with test scores that you created in number 5. 

```{r}

Scores
ifelse(Scores>50, print("TRUE"),print("FALSE"))

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: `ifelse` worked.
